---
title: "Final Project"
author: "Sandeep Nair"
date: "2/25/2022"
output: output= github_document
---

## Abstract

<Short summary of the report. Go over each section briefly>

## Introduction

COVID 19 in an ongoing pandemic which was first identified in December 2019. In its early stages, epidemiology experts emphasized the importance of collecting data on policy decisions. As a result, large, robust data sets are made available on the World heath Organisations Website.

My group has identified two potential questions of interest for this report:

-   We want to find out if two countries with similar populations and land area have a similar or different average new Covid cases rate through the timeline and similar or different average death rates. We chose United Kingdom and France for this comparison.

-   To check if there is a correlation between the infected rate and Response policy. Stringency index ,which is calculated by the Oxford COVID-19 Response team, is used to numerical represent response policy.

-   Additionally, we would identify two countries with contrasting policy response (stringent vs relaxed) to check if policy response had a significant effect on disease control.

## Background

The data obtained from WHO includes the date reported, country code, country, WHO region, new Covid cases per date, cumulative Covid cases, new death rates, and cumulative deaths from dates 03-Jan-2020 to 17-Feb-2022. Although our data has been presented to us through the World Health Organiztion, these numbers are subject to bias based on the number of testing a country does and if the country is truly giving correct numbers to the WHO.

The stringency data set was obtained from <https://ourworldindata.org/coronavirus>. The stringency index is calculated based on nine response indicators including school closures, workplace closures, and travel bans, rescaled to a value from 0 to 100 (100 = strictest). This index is calculated by Oxford COVID-19 Government Response Tracker (oxCGRT).

The two dataset use different country codes. To join the two datasets, we need to convert the three letter country code in the stringency dataset to two letters. We need a database which matches two letter codes to three letter codes. This was obtained as a .csv file from <https://gist.github.com/tadast/8827699#file-countries_codes_and_coordinates-csv-L1>.

Code for data wrangling and joining:

```{r message = FALSE}
library(tidyverse)
library(reshape2)
```

```{r message = FALSE}
Covid <- read_csv("./data/WHO-COVID-19-global-data.csv")
head(Covid)
stringency = read_csv("./data/covid-stringency-index.csv")
head(stringency)
```

```{r}
country = read_csv("./data/countries_codes_and_coordinates.csv")
colnames(country)[3] = "Code"
```

```{r}
stringency_2letter = left_join(stringency, country, by="Code")
stringency_2letter = stringency_2letter[,c(3,6,4)]
colnames(stringency_2letter)[2] = "Code_2"
```

```{r}
joined = inner_join(Covid, stringency_2letter, by = c("Date_reported" = "Day","Country_code" = "Code_2" ))
head(joined)

```

## Descriptive Analysis (by 2/25)

Tasks

1.  Identify countries with severe restrictions and relaxed policies for case study
```{r}
japan= joined %>% filter(Country_code =="JP" )
mean((japan$stringency_index))
Canada = joined %>% filter(Country_code == "CA")
mean(Canada$stringency_index)
t.test(japan$stringency_index, Canada$stringency_index)
```

From the t-Test we can see that there is a statistically significant difference between the stringency indices of Canada and Japan. We can say with 95 confidence that Canada's stringency measure was 24.916 to 21.99 units higher than Japan's.

2.  Plot of stringency index overtime in selected countries
```{r}
string_combined = inner_join(japan, Canada, by= "Date_reported")
string_combined = string_combined[,c(1,9,17)]
colnames(string_combined) = c("Date","JP","CAN")
string_melt = melt(string_combined, id.vars = "Date", variable.name = "country")
colnames(string_melt)[3]= "Stringency_index"
ggplot(data = string_melt, aes(Date, Stringency_index))+geom_line(aes(colour = country))

```


3.  Box plot comparing infection levels.

## Inferential Analysis(by 2/27)

Tasks

1.  Select either Anova or linear regression to model data

2.  Explain selected model, parameters, assumptions

3.  Hypothesis testing and conclusion

## Sensitivity Analysis (by 2/28)

Task - Examine the residual plot of the fitted model. we need to explain your findings in these plot

## Discussion

Tasks -

1.  A brief recap of this project.

2.  Findings in the inferential analysis interpreted in the context of WHO findings

3.  Suggestions for future research and/or policy making given your findings.

4.  Caveats of the current analysis

## Acknowledgement

Discussed project with my group member, Oh Seung Kwon

## Reference 

1.  <https://ourworldindata.org/coronavirus>.

2.   <https://gist.github.com/tadast/8827699#file-countries_codes_and_coordinates-csv-L1>.

<update as we go>